In-Band Synchronization for\\Distributed SDN Control Planes\thanks{Supported by the European Institute of Innovation \& Technology (EIT) project 13153 (Software Defined Networking) and the FP7 EU project UNIFY.}

### 2015-11-16 ccr-camera/inband-sync.tex
release tokens

\vspace{1mm}
\noindent\textbf{Note on efficiency.}

\stefan{We conclude by observing that our implementation is switch space efficient,
in the sense that it requires only a constant number of flow entries per memory address
and resources scale linearly in the number of controllers. Also, in case of controller failures,
at most a constant amount of space is lost.}
\petr{Unclear which ``resources'' are meant here? Should this
  discussion be postponed to policy updates, where ids indeed grow
  linearly with the number of updates?}

\stefan{We also note that while we described our scheme in a simplified
way where numbers continuously increase, it is straight-forward to use a modulo
scheme. While in our case 32-bit numbers are used, 
the flow table width (number of bits which can be matched in parallel)
of modern switches is 40-bits and more.}
 \liron{I think we should say something like: "... of modern switches can get as high as a few hundreds'. }

@@

Let us now revisit our load-balancing application
discssed in Section~\ref{sec:motivation}.
A simple strategy to balance load in our 2-link example
 (Figure~\ref{fig:CAS-example}) fails in the presence of
multiple controllers, even if 
bundling and $\checko$ features are used:
the flows are defined over ``independent'' 
flow spaces.

As we show in the following, our mechanisms can be used 
to render such algorithms concurrent.
In particular, we can implement a
generic template to
implement concurrent policy compositions of a user-specific
$\textsc{\ufunc}()$ function~\cite{stn}.
This way, for example, by choosing our load-balancing
function $\textsc{lb\_update}$, 
the pathological scenario described in 
Figure~\ref{fig:CAS-example} can be resolved.
% if we put
%the policy update in our transactional $\execatomic$ construct together
% with a \emph{Compare-and-Set} (CAS) operation on the policy identifier.

@@

Algorithm~\ref{alg:update} is a more efficient alternative 
to Algorithm~\ref{alg:simple-update}. It uses our id-claimer abstraction
to bound the number of used policy identifiers.
Here, the controller first reads the current policy id and claims to prevent another controller from using it 
for a
different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}).
Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
Finally, within an atomic transaction (\textbf{\execatomic}), it verifies whether the chosen policy id has not
been claimed and that
the policy has not been concurrently modified, and then updates the policy.

@@

To investigate the feasibility of our approach,
we implemented a proof-of-concept of our algorithms
as a python library which wraps ovs-ofctl, 
a well-known command line tool for monitoring and administering OpenFlow switches. 
\stefan{The code is made publicly available at TODO LIRON.}\liron{https://github.com/lironsc/of-sync-lib}

@@

For instance, we experimented with a setting where three processes
(controllers) issue thousands of concurrent policy updates
to the switch, where 
each update consists in replacing a single rule with another one.
When verifying the transactions, we could confirm that 
despite the concurrency, 100 \% of the transactions have been successfully completed.
Of course, due to the concurrency, multiple transactions also needed to abort,
due to conflicts.

While our framework can be used as synchronization tool itself, we 
currently also work on implementing it as an integral part of existing controllers 
such as Ryu.

@@

We believe that our centralized in-band
synchronization approach is natural and attractive, as it avoids the
complexities and limitations of distributed out-of-band agreement protocols, which may 
introduce additional points of potential failures.
Moreover, the switch is the element implementing the
controller policies anyway.
Interestingly, the mechanisms presented in this paper are simple and can be implemented using the \emph{standard OpenFlow}
protocol: they do not require any protocol/hardware extensions as postulated in recent literature~\cite{stn,netpaxos}.
Our work may hence also contribute to the ongoing discussion of what can be implemented
in-band in today's OpenFlow protocol~\cite{reclaim},
as well as to useful high-level concurrency objects and
language abstractions~\cite{pyretic}.
In general, we understand our paper as a first step, and more
research is required to investigate and compare useful
inband synchronization abstractions, also in terms of performance.
{
%\begin{footnotesize}
\bibliographystyle{abbrv}
\bibliography{references}  % main.bib is the name of the Bibliography in this case
%\end{footnotesize}
}

>>>>

@@

Let us now revisit our load-balancing application
discussed in Section~\ref{sec:motivation}.
A simple strategy to balance load in our 2-link example
 (Figure~\ref{fig:CAS-example}) fails in the presence of
multiple controllers, even if 
bundling and $\checko$ features are used:
the flows are defined over ``independent'' 
flow spaces.

%[[PK ``render concurrent''? we already discuss a concurrent setting...
%As we show in the following, our mechanisms can be used 
%to render such algorithms concurrent.
Using CAS, we can implement a
generic template 
for concurrent policy composition via a user-specific
$\textsc{\ufunc}()$ function~\cite{stn}.
This can be, for example, a specific load-balancing
function $\textsc{lb\_update}$, %\petr{Did we talk about this function before?}
which, when applied atomically, excludes the pathological scenario described in 
Figure~\ref{fig:CAS-example}. 
We use CAS to exactly achieve this atomicity.
% if we put
%the policy update in our transactional $\execatomic$ construct together
% with a \emph{Compare-and-Set} (CAS) operation on the policy identifier.

@@

A less trivial Algorithm~\ref{alg:update} adresses the problem of
policy identifiers growing without bound in  
Algorithm~\ref{alg:simple-update} by, additionally, using our id-claimer abstraction.
%to bound the number of used policy identifiers.
Here, the controller first reads the current policy id and 
claims it, which prevents another controller from using the id 
for a different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}).Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
Finally, within an atomic transaction (\textbf{\execatomic}), it verifies whether the chosen policy id has not
been claimed and that
the policy has not been concurrently modified, and then updates the
policy.
Unlike Algorithm~\ref{alg:simple-update}, this algorithm is only
\emph{lock-free}, as it allows scenarios in which only one live controller
succeeds in installing its updates. We can easily make it
wait-free by

\vspace{1mm}
\noindent\textbf{A note on efficiency.}

We observe that our implementations of synchronization primitives
(Algorithms~\ref{alg:claim}-\ref{alg:compare}) 
%is \emph{switch-space efficient}
%in the sense that 
require only a constant number of flow entries per memory address.
The number of used addresses is constant for  
Algorithm~\ref{alg:simple-update} and linear in the number of
controllers for Algorithm~\ref{alg:update}. 
%[[PK not constant: proportional to the number of failures. But this
%is kind of obvious.
%Also, in case of controller failures, at most a constant amount of space is lost.
%]]

\stefan{We also note that while we described our scheme in a simplified
way where numbers continuously increase, it is straight-forward to use a modulo
scheme. While in our case 32-bit numbers are used, 
the flow table width (number of bits which can be matched in parallel)
of modern switches is 40-bits and more.}

 \liron{I think we should say something like: "... of modern switches can get as high as a few hundreds'. }

\petr{The discussion is very confusing. Do we need it?}

@@

To investigate the feasibility of our approach,
we implemented a proof-of-concept of our algorithms
as a python library which wraps ovs-ofctl, 
a well-known command line tool for monitoring and administering OpenFlow switches. 
The code is publicly available at \url{https://github.com/lironsc/of-sync-lib}.

@@

For instance, we experimented with a setting where three processes
(controllers) issue thousands of concurrent policy updates
to the switch, 
each update consisting in replacing a single rule with another one.
%[[PK unclear
%When verifying the transactions, 
%]]
We could confirm that 
despite the concurrency, 100 \% of the updates have been
eventually completed.
Of course, due to the concurrency, some transactions also needed to abort
because of conflicts.

While our framework can be used as a synchronization tool itself, we 
currently also work on implementing it as an integral part of existing controllers 
such as Ryu. 
\petr{Did not get the difference between ``itself'' and ``part of
  existing controllers''.}

@@

We believe that our centralized in-band
synchronization approach is natural and attractive, as it avoids the
complexities and limitations of distributed out-of-band agreement protocols, which may 
introduce additional points of potential failures.
Moreover, the switch is the element implementing the controller policies anyway.
The mechanisms presented in this paper are simple and can be implemented using the \emph{standard OpenFlow}
protocol: they do not require any protocol/hardware extensions as postulated in recent literature~\cite{stn,netpaxos}.
Our work may hence also contribute to the ongoing discussion of what can be implemented
in-band in today's OpenFlow protocol~\cite{reclaim},
as well as to useful high-level concurrency objects and
language abstractions~\cite{pyretic}.
In general, we see our paper as a first step, and more
work is required to investigate and compare useful
in-band synchronization abstractions, also in terms of performance.

{
%\begin{footnotesize}
\bibliographystyle{abbrv}
\bibliography{references}  % main.bib is the name of the Bibliography in this case
%\end{footnotesize}
}
